---
title: "Interactive map"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    source: embed
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(leaflet)
library(plotly)
library(flexdashboard)
library(shiny)
```

```{r}
clean_df = read_csv("./data_shiny/cleaned.csv") %>% 
  mutate(
    college_attendance_status = recode(college_attendance_status, "Yes" = "1", "No" = "0"),
    college_attendance_status = as.numeric(college_attendance_status)
    ) %>% 
  mutate(
    metropolitan_status = recode(metropolitan_status, "in metropolitian but mixed of central city" = "In metropolitan", "in metropolitan and central city" = "In metropolitan", "in metropolitian but not in central city" = "In metropolitan")
  )
clean_all1 = 
  clean_df %>% 
  mutate(states = recode(states, .default = "ALL"))
clean_all2 = 
  clean_df %>% 
  mutate(race = recode(race, .default = "ALL"))
clean_all3 = 
  clean_df %>% 
  mutate(race = recode(race, .default = "ALL"),
         states = recode(states, .default = "ALL")
         )

clean_final = 
  rbind(clean_all1, clean_all3, clean_df, clean_all2)
```

College Attendance Rate Over Time
===================================== 

Column{.sidebar}
-----------------------------------------------------------------------


```{r for_sidebar}
states_choices = clean_final %>% pull(states) %>% unique()
metropolitan_choices = clean_final %>% pull(metropolitan_status) %>% unique()
children_range = clean_final %>% pull(number_of_children)
income_range = clean_final %>% pull(number_of_children)
race_choice = clean_final %>% pull(race) %>% unique()

selectInput(
  "states_choice",
  label = h3("Select State"),
  choices = states_choices,
  selected = "NY"
)

radioButtons(
  "race_choice",
  label = h3("Select Race"),
  choices = race_choice,
  width = "160%",
  selected = "ALL"
)
sliderInput(
  "income_range",
  h3("Annual Family Income"),
  -20000, 150000, 
  value = c(-20000, 150000))

checkboxGroupInput(
  "metropolitan_choice",
  h3("Metropolitan Status"),
  width = "160%",
  choices = metropolitan_choices,
  selected = c("mixed of metropolitian status", "In metropolitan", "Not in metropolitan area")
)
```

Column {.tabset}
-----------------------------------------------------------------------

### College Attendance Rate Cross Time

```{r}
renderPlotly({
clean_final %>% 
  filter(
      states == input[["states_choice"]],
      metropolitan_status == input[["metropolitan_choice"]],
      anually_family_income %in% input[["income_range"]][1]:input[["income_range"]][2],
      race == input[["race_choice"]]) %>%
  group_by(year) %>%
  mutate(
    number_1 = "1",
    number_1 = as.numeric(number_1),
    college_attendance_rate = sum(college_attendance_status)/sum(number_1)
  ) %>% 
    ungroup() %>% 
  plot_ly(
    y = ~college_attendance_rate, x = ~year, color = ~states, mode = "lines+markers",
    type = "scatter", colors = "viridis") %>% 
    layout(
      yaxis = list(
        range = c(0,1)
      ),
      title = "College Attendance Rate Over Time (2008-2018)"
    )
})
```

### Description  

The line plot illustrates the college attendance rate from 2008 to 2018 across 52 states in the United States and the college attendance rate for different annual family income, different metropolitan status, and different races.  

**Select ALL to find more information about the college attendance rate**


College Attendance Rate Map
===================================== 

Column {.sidebar}
-----------------------------------------------------------------------


```{r for_sidebar2}
year_choices = clean_final %>% pull(year) %>% unique()
# year select box 
selectInput("year_choices", label = h3("Choose year"),
						choices = year_choices,
						selected = 2008)
```


Column {.tabset}
-----------------------------------------------------------------------

### US MAP  

```{r us_map}
# for convert state abbreviations to state names
state_abb = state.name
names(state_abb) = state.abb

#clean data for map
clean_df_final = 
        clean_df %>% 
        select(year, states, college_attendance_status)
# plot_geo
renderPlotly({
	# map plot setting
	geo1 <- list(
	scope = "usa",
	projection = list(type = "state"),
	showlakes = TRUE,
	lakecolor = toRGB("white")
	)
	
	# display value using color intensity
	clean_df_final %>% 
	filter(year == input$year_choices) %>%
	group_by(year,states) %>%
  mutate(
    number_1 = "1",
    number_1 = as.numeric(number_1),
    college_attendance_rate = sum(college_attendance_status)/sum(number_1)
  ) %>% 
    ungroup() %>% 
	plot_geo(locationmode = "USA-states") %>% 
	add_trace(
		z = ~ college_attendance_rate,
		locations = ~states,
		color = ~ college_attendance_rate,
		colors = "Reds"
		) %>%
	layout(
		geo = geo1,
		title = "College Attendance Rate Map",
		legend = list(x = 100, y = 0.5)
	)
})
```

### Description  

The College Attendance Rate map allows users to toggle between different years (2008-2018), visualizing regional differences across the U.S.  
