---
title: "Interactive map"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    source: embed
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(leaflet)
library(plotly)
library(flexdashboard)
library(shiny)
```

```{r}
clean_df = read_csv("./cleaned.csv") %>% 
  mutate(
    college_attendance_status = recode(college_attendance_status, "Yes" = "1", "No" = "0"),
    college_attendance_status = as.numeric(college_attendance_status)
    ) %>% 
  mutate(
    metropolitan_status = recode(metropolitan_status, "in metropolitian but mixed of central city" = "In metropolitan", "in metropolitan and central city" = "In metropolitan", "in metropolitian but not in central city" = "In metropolitan")
  )
clean_all1 = 
  clean_df %>% 
  mutate(states = recode(states, .default = "ALL"))
clean_all2 = 
  clean_df %>% 
  mutate(metropolitan_status = recode(metropolitan_status, .default = "ALL"))
clean_all3 = 
  clean_df %>% 
  mutate(metropolitan_status = recode(metropolitan_status, .default = "ALL"),
         states = recode(states, .default = "ALL")
         )
clean_final = 
  rbind(clean_all1, clean_all3, clean_df,clean_all2)
```

College Attendance Rate Over Time
===================================== 

Column{.sidebar}
-----------------------------------------------------------------------


```{r for_sidebar}
states_choices = clean_final %>% pull(states) %>% unique()
metropolitan_choices = clean_final %>% pull(metropolitan_status) %>% unique()
children_range = clean_final %>% pull(number_of_children)
income_range = clean_final %>% pull(number_of_children)

selectInput(
  "states_choice",
  label = h3("Select State"),
  choices = states_choices,
  selected = "NY"
)

sliderInput(
  "income_range",
  h3("Annual Family Income"),
  -20000, 150000, 
  value = c(-20000, 150000))

radioButtons(
  "metropolitan_choice",
  h3("Metropolitan Status"),
  choices = metropolitan_choices
)
```

Column {.tabset}
-----------------------------------------------------------------------

### College Attendance Rate Cross Time

```{r}
renderPlotly({
clean_final %>% 
  filter(
      states == input[["states_choice"]],
      metropolitan_status == input[["metropolitan_choice"]],
      anually_family_income %in% input[["income_range"]][1]:input[["income_range"]][2]) %>%
  group_by(year) %>%
  mutate(
    number_1 = "1",
    number_1 = as.numeric(number_1),
    college_attendance_rate = sum(college_attendance_status)/sum(number_1)
  ) %>% 
    ungroup() %>% 
  plot_ly(
    y = ~college_attendance_rate, x = ~year, color = ~states, mode = "lines+markers",
    type = "scatter", colors = "viridis") %>% 
    layout(
      yaxis = list(
        range = c(0,1)
      ),
      title = "College Attendance Rate Over Time (2008-2018)"
    )
})
```

### Description  

The line plot illustrates the college attendance rate from 2008 to 2018 across 52 states in the United States and the college attendance rate for different annual family income and different metropolitan status.  

**Select ALL to find more information about the college attendance rate**


College Attendance Rate Map
===================================== 

Column {.sidebar}
-----------------------------------------------------------------------


```{r for_sidebar2}
year_choices = clean_final %>% pull(year) %>% unique()
# year select box 
selectInput("year_choices", label = h3("Choose year"),
						choices = year_choices,
						selected = 2008)
```


Column {.tabset}
-----------------------------------------------------------------------

### US MAP  

```{r us_map}
# for convert state abbreviations to state names
state_abb = state.name
names(state_abb) = state.abb

# plot_geo
renderPlotly({
	# map plot setting
	geo1 <- list(
	scope = "usa",
	projection = list(type = "state"),
	showlakes = TRUE,
	lakecolor = toRGB("white")
	)
	
	# display value using color intensity
	clean_df %>% 
	filter(year == input[["year_choices"]]) %>%
	group_by(year) %>%
  mutate(
    number_1 = "1",
    number_1 = as.numeric(number_1),
    college_attendance_rate = sum(college_attendance_status)/sum(number_1)
  ) %>% 
    ungroup() %>% 
	plot_geo(locationmode = "USA-states") %>% 
	add_trace(
		z = ~college_attendance_rate,
		locations = ~states,
		color = ~college_attendance_rate,
		colors = "Reds",
		# add text
		text = ~paste(state_abb[states], sep = "")
		) %>%
	layout(
		geo = geo1,
		title = "US Map - State Statistics",
		legend = list(x = 100, y = 0.5)
	)
})
```

### Description  
